### 2015-04-08 prelim.tex
Bringing everything up to date

\section{Preliminaries}\label{sec:pois_review}
\subsection{Poisson Surface Reconstruction}
We denote the indicator function of a model $M$ as $$ \chi_M(\mathbf{x}) :=
\begin{cases}
1, & \text{if }\mathbf{x}\text{ is in the interior of the model } M \\
\frac{1}{2}, & \text{if } \mathbf{x}\text{ is on the model's surface}\\
0, & \text{if }\mathbf{x}\text{ is exterior to } M
\end{cases}. $$
Integral to formulating surface reconstruction as a Poisson problem, is the observation that if one obtains some function $\vec{V}(\mathbf{x})\approx{\nabla}\chi_M$ that approximates the smoothed gradient of the indicator function, finding an approximation to the indicator function reduces to finding a $\inda$ such that {\small 
\begin{equation} \label{eq:poisson}
	\lapl{\inda} = {\grad} \cdot \vec{V}.
\end{equation}}
We define the set of input surface points to be $P:=\{(\mathbf{p}_1,\mathbf{n}_1), (\mathbf{p}_2,\mathbf{n}_2) \hdots (\mathbf{p}_M,\mathbf{n}_M) : (\mathbf{p}_i,\mathbf{n}_i) \in
\realn^3 \times \realn^3 \}$ where $\mathbf{p}_i$ and $\mathbf{n}_i$ are the position and normal vector of the $i^{th}$ sample respectively. Orientation can be succinctly represented with only two values, however, it is convenient to keep it as the normal at a point, because the length of the normal can be used to encode the sampling density of the point-set at a point. For the following discussion, we also need to introduce the notation $\patch_{\mathbf{s}}$, which denotes a patch of the surface model about the sample location $\mathbf{s}$.

@@

Typically, as memory is a limited quantity, we are only interested in a finite number of points on a lattice. We denote the set of such points as $\{\mathbf{s}_1,\mathbf{s}_2,\ \hdots  \ \mathbf{s}_N\} \subset \mathbf{L}\intn^3.$ Corresponding to each lattice site is a coefficient $c_k$; we collect all such coefficients in the column vector $\mathbf{c}:=\left[c_1 \ c_2 \ \hdots \ c_N \right]^T.$ We also use the short hand notation $\varphi_k(\vx) := \varphi(\frac{\vx}{h}-\mathbf{s}_k)$; from this perspective, our function space is now {\small 
\begin{equation}
	\left\{ \displaystyle\sum_{k=0}^N c_k \varphi_k(\vx) \right\} \subset \fspce{\varphi}{\lattice{L}_h}.
\end{equation}}
Further, we denote $\mathbf{\Phii}$ to be the $M \times N$ matrix whose elements are given by sampling the basis functions anchored at each lattice site, or $\Phii_{i,j} := \varphi_j(\vx_i)$. Clearly, if our data are sampled at the lattice sites ($\vx_i = \mathbf{s}_i$), then the problem of finding an interpolative scheme reduces to finding $\mathbf{c}$ such that $\mathbf{\Phii c=f},$ which only involves inverting the square matrix $\mathbf{\Phii}.$ However, our sample locations are almost certain to not coincide with lattice sites and we must consider alternative schemes for finding $\mathbf{c}$, which we'll come back to in \SC{sec:vari_review}

@@

In order to ensure that any function $g \in \aspace{\lattice{L}}{\varphi}$ has a unique representation in terms of a finite energy coefficient sequence $c[\cdot]$, the lattice translates $\{\varphi(\cdot - \mat{L}\vect{n})\}_{\vect{n}  \in \ints^s}$ must form a Riesz basis~\cite{kreyszig89}, i.e. there exist positive constants $a$ and $b$ such that $\forall c[\cdot] \in l_2(\ints^s)$,
\begin{equation*}
  a \norm{c}^2_{l_2(\ints^s)} \le \Ltwonormbig{\sum_{\vect{n} \in \ints^s}
  c[\vect{n}] \varphi(\vect{x} - \mat{L}\vect{n})}^2 \le b
	\norm{c}^2_{l_2(\ints^s)}.
\end{equation*}

@@

If the generating function that spans the space satisfies the \emph{Strang-Fix} conditions of order $k$, and $f$ belongs to the Sobolev space of at least order $k$, then it can be shown that the minimum approximation error (in the least-squares sense) decays asymptotically as $O(h^k)$, and the space is said to have approximation order $k$. We often only have access to $f$ through a discrete amount of samples, $\mathbf{f} := \left[ f_1 \ f_2 \ \hdots \ f_M \right]^T$ located at sample points $S := \{\vx_1, \vx_2 \hdots \vx_M \}$ respectively. When this is the case, the minimum-error approximation is unattainable. However, interpolative and quasi-interpolative methods \cite{quasibcc,interprev}, which can be seen as oblique projections onto $\fspce{\varphi}{h\mathbf{L}}$, maintain the same order of accuracy.

@@

\subsection{Box Splines} \label{sec:box_review}
Box splines are compact piece-wise polynomial functions $ M_\mathbf{\Xii} : \realn^s \to \realn.$ The matrix $\mathbf{\Xii}$ denotes a collection of direction vectors, $\mathbf{\Xii} := \left[ \xi_1 \ \xi_2 \hdots \xi_n\right].$ When $n=s$, the function $M_\mathbf{\Xii}(\vx)$ is defined to be $\frac{1}{det|\mathbf{\Xii}|}$ inside the polytope formed by the Minkowski sum of the direction vectors of $\mathbf{\Xii}$, and zero otherwise (that is, $\frac{1}{det|\mathbf{\Xii}|}$ when $\vx \in \{ \mathbf{\xi}_1 t_1 + \hdots + \mathbf{\xi}_s t_s  : t_i\in [0,1) \})$. When $n > s$, the box spline is recursively defined as {\small 
\begin{equation} \label{eq:box_def}
	M_{\mathbf{\Xii}}(\vx) := \int_0^1{\!M_{\mathbf{\Xii/\xi}_n}(\vx - t\mathbf{\xi}_n)} \mathrm{d}t,
\end{equation}}
where $\!M_{\mathbf{\Xii/\xi}_n}$ is the matrix obtained by removing the column vector $\xi_n$ from $\mathbf{\Xii}$. Box splines are a natural fit for our purposes, and are chosen to span our function spaces. One motivating factor that encourages the use of box splines on the BCC lattice over the Cartesian lattice is that the support of the box splines on the BCC lattice contains less lattice sites as opposed to their Cartesian counterparts. At the same time, they maintain the same order of accuracy and smoothness. This implies that one should expect roughly the same quality reconstruction, with less memory accesses per reconstructed value.

@@

\subsection{Irregular Sampling}
\label{sec:vari_review}
%When reconstructing functions whose data are sampled regularly on a lattice, finding an oblique projection of the original function in the target function space is relatively straight-forward. However, except on very rare occasions, samples placed over the surface of a model do not coincide with lattice sites. A solution to this problem is to ``re-sample'' the input point set onto said lattices sites.
In the work by Kazdhan et al. \cite{fftk} the range data are splatted via a linear filter onto a regular Cartesian grid. This tends to over-smooth the resulting reconstruction, as it allows the surface to deviate from the original sample locations. Ideally, we would like to find a function that not only interpolates the input data, but provides the representation with the smallest possible ``shell'' about the initial point set (i.e., the fewest non zero coefficients) while still interpolating the data and eliciting a unique solution.

Thus, we recast the problem of finding a representation of $g$ within our function space as an optimization problem. There are similar works that impose variational constraints on random point samples for images \cite{variational}, as well as multi-resolution techniques for volumetric data \cite{onvari} and an extension to box spline spaces \cite{xu2012rec}. In the same vein, we seek a function $g \in \fspce{\varphi}{h\mathbf{L}}$ that attempts to interpolate the input data, and minimize a functional $E$. Specifically, we seek a $g$ that minimizes {\small 
\begin{equation} \label{eq:minexp}
 	\sump_{i=1}^{M} (g(\vx_i) -  f_i)^2 + E(g).
\end{equation}}
It is important to introduce a bit of notation before proceeding, $\dv{\mathbf{v}} f := \mathbf{v} \cdot \grad f$ defines the directional derivative of $f$ in the direction $\mathbf{v}$. The Beppo-Levi inner product of order $n$ is defined as $\blip{n}{f}{g}:= \sum_{|\mathbf{p}|=n}({n!}/{\mathbf{p}!}) \innerp{\mathbf{D^p}f}{\mathbf{D^p}g}$, where $\mathbf{p}$ is an integer vector such that $\left| \mathbf{p}\right| := p_1 + p_2 + p_3 $. We also use the short hand $\mathbf{D^p}f := \partial^{\left| \mathbf{p}\right|}f/({\partial^{p_1}x_1}{\partial^{p_2}x_2}{\partial^{p_3}x_3}).$ The second order Beppo-Levi norm of $f$ is defined as $\bln{2}{f}^2:=\blip{2}{f}{f}$.

@@

Finding $g$ that minimizes equation \EQ{eq:minexp} involves rewriting the minimization problem in terms of $\mathbf{c}$ \cite{xu2012rec}. Following some straightforward algebraic manipulations (see \cite{xu2012rec} for details), the coefficient vector can be obtained by taking $\mathbf{c} = ( \mathbf{\Phii}^T\mathbf{\Phii} + \lambda_1\mathbf{G} + \lambda_2\mathbf{S})^{-1}\mathbf{\Phii}^T\mathbf{f}$, where $G_{i,j}:=\innerp{\varphi_i}{\varphi_j}$ and $S_{i,j}:=\blip{2}{\varphi_i}{\varphi_j}$. Since $\varphi$ is known in closed form, the entries of both $\mathbf{G}$ and $\mathbf{S}$ can be evaluated analytically. The above vector equation can be efficiently solved using the Conjugate Gradient method.
%\begin{equation}
%	\mathbf{c} = \left( \mathbf{\Phii}^T\mathbf{\Phii} + \lambda_1\mathbf{G} + \lambda_2\mathbf{S}\right)^{-1}\mathbf{\Phii}^T\mathbf{f},
%\label{eq:variationalSolution}
%\end{equation}

>>>>

\section{Preliminaries}\label{sec:pois_review}
\subsection{Poisson Surface Reconstruction}
We denote the indicator function of a model $M$ as $$ \chi_M(\mathbf{x}) :=
\begin{cases}
1, & \text{if }\mathbf{x}\text{ is in the interior of the model } M \\
\frac{1}{2}, & \text{if } \mathbf{x}\text{ is on the model's surface}\\
0, & \text{if }\mathbf{x}\text{ is exterior to } M
\end{cases}. $$
The key observation of the Poisson surface reconstruction formulation, is that if one obtains a function $\vec{V}(\mathbf{x})$ with $\vec{V}(\mathbf{x})\approx{\nabla}\chi_M$, that approximates the smoothed gradient of the indicator function, finding an approximation to the indicator function reduces to finding a $\inda$ such that {\small 
\begin{equation} \label{eq:poisson}
	\lapl{\inda} = {\grad} \cdot \vec{V}.
\end{equation}}
We define the set of input surface points to be $P:=\{(\mathbf{p}_1,\mathbf{n}_1), (\mathbf{p}_2,\mathbf{n}_2) \hdots (\mathbf{p}_M,\mathbf{n}_M) : (\mathbf{p}_i,\mathbf{n}_i) \in
\realn^3 \times \realn^3 \}$ where $\mathbf{p}_i$ and $\mathbf{n}_i$ are the position and normal vector of the $i^{th}$ sample respectively. Orientation can be succinctly represented with only two values, however, it is convenient to keep it as the normal at a point, because the length of the normal can be used to encode the sampling density of the point-set at a point. For the following discussion, we also need to introduce the notation $\patch_{\mathbf{s}}$, which denotes a patch of the surface model about the sample location $\mathbf{s}$.

@@

Typically, as memory is a limited quantity, we are only interested in a finite number of points on a lattice. We denote the set of such points as $\{\mathbf{s}_1,\mathbf{s}_2,\ \hdots  \ \mathbf{s}_N\} \subset \mathbf{L}\intn^3.$ Corresponding to each lattice site is a coefficient $c_k$; we collect all such coefficients in the column vector $\mathbf{c}:=\left[c_1 \ c_2 \ \hdots \ c_N \right]^T.$ We also use the short hand notation $\varphi_k(\vx) := \varphi(\frac{\vx}{h}-\mathbf{s}_k)$; from this perspective, our function space is now {\small 
\begin{equation}
	\left\{ \displaystyle\sum_{k=0}^N c_k \varphi_k(\vx) \right\} \subset \fspce{\varphi}{\lattice{L}_h}.
\end{equation}}
Further, we denote $\mathbf{\Phii}$ to be the $M \times N$ matrix whose elements are given by sampling the basis functions anchored at each lattice site, or $P_{i,j} := \varphi_j(\vx_i)$. Clearly, if our data are sampled at the lattice sites ($\vx_i = \mathbf{s}_i$), then the problem of finding an interpolative scheme reduces to finding $\mathbf{c}$ such that $\mathbf{\Phii c=f},$ which only involves inverting the square matrix $\mathbf{\Phii}.$ However, our sample locations are almost certain to not coincide with lattice sites and we must consider alternative schemes for finding $\mathbf{c}$, which we'll come back to in \SC{sec:vari_review}

@@

In order to ensure that any function $g \in \aspace{\lattice{L}_h}{\varphi}$ has a unique representation in terms of a finite energy coefficient sequence $c[\cdot]$, the lattice translates $\{\varphi(\cdot - \mat{L}\vect{n})\}_{\vect{n}  \in \ints^s}$ must form a Riesz basis~\cite{kreyszig89}, i.e. there exist positive constants $a$ and $b$ such that $\forall c[\cdot] \in l_2(\ints^s)$,
\begin{equation*}
  a \norm{c}^2_{l_2(\ints^s)} \le \Ltwonormbig{\sum_{\vect{n} \in \ints^s}
  c[\vect{n}] \varphi(\vect{x} - \mat{L}\vect{n})}^2 \le b
	\norm{c}^2_{l_2(\ints^s)}.
\end{equation*}

@@

If the generating function that spans the space satisfies the \emph{Strang-Fix} conditions of order $k$, and $f$ belongs to the Sobolev space of at least order $k$, then it can be shown that the minimum approximation error (in the least-squares sense) decays asymptotically as $O(h^k)$, and the space is said to have approximation order $k$. We often only have access to $f$ through a discrete amount of samples, $\mathbf{f} := \left[ f_1 \ f_2 \ \hdots \ f_M \right]^T$ located at sample points $S := \{\vx_1, \vx_2 \hdots \vx_M \}$ respectively. When this is the case, the minimum-error approximation is unattainable. However, interpolative and quasi-interpolative methods \cite{quasibcc,interprev}, which can be seen as oblique projections onto $\aspace{\lattice{L}_h}{\varphi}$, maintain the same order of accuracy.

@@

\subsection{Box Splines} \label{sec:box_review}
Box splines are compact piece-wise polynomial functions $ M_\mathbf{\Xii} : \realn^s \to \realn.$ The matrix $\mathbf{\Xii}$ denotes a collection of direction vectors, $\mathbf{\Xii} := \left[ \xi_1 \ \xi_2 \hdots \xi_n\right].$ When $n=s$, the function $M_\mathbf{\Xii}(\vx)$ is defined to be ${1}/|{\det\mathbf{\Xii}}|$ inside the polytope formed by the Minkowski sum of the direction vectors of $\mathbf{\Xii}$, and zero otherwise (that is, ${1}/{|\det\mathbf{\Xii}|}$ when $\vx \in \{ \mathbf{\xi}_1 t_1 + \hdots + \mathbf{\xi}_s t_s  : t_i\in [0,1) \})$. When $n > s$, the box spline is recursively defined as {\small 
\begin{equation} \label{eq:box_def}
	M_{\mathbf{\Xii}}(\vx) := \int_0^1{\!M_{\mathbf{\Xii/\xi}_n}(\vx - t\mathbf{\xi}_n)} \mathrm{d}t,
\end{equation}}
where ${\mathbf{\Xii/\xi}_n}$ is the matrix obtained by removing the column vector $\xi_n$ from $\mathbf{\Xii}$. Box splines are a natural fit for our purposes, and are chosen to span our function spaces. One motivating factor that encourages the use of box splines on the BCC lattice over the Cartesian lattice is that the support of the box splines on the BCC lattice contains less lattice sites as opposed to their Cartesian counterparts. At the same time, they maintain the same order of accuracy and smoothness. This implies that one should expect roughly the same quality reconstruction, with less memory accesses per reconstructed value.

@@

\subsection{Irregular Sampling}
\label{sec:vari_review}
%When reconstructing functions whose data are sampled regularly on a lattice, finding an oblique projection of the original function in the target function space is relatively straight-forward. However, except on very rare occasions, samples placed over the surface of a model do not coincide with lattice sites. A solution to this problem is to ``re-sample'' the input point set onto said lattices sites.
In the work by Kazdhan et al. \cite{fftk} the range data, $P$, are splatted via a linear filter onto a regular Cartesian grid. This tends to over-smooth the resulting reconstruction, as it allows the surface to deviate from the original sample locations. Ideally, we would like to find a function that not only interpolates the input data, but provides the representation with the smallest possible ``shell'' about the initial point set (i.e., the fewest non zero coefficients) while still interpolating the data and eliciting a unique solution.

Thus, we recast the problem of finding a representation of $g$ within our function space as an optimization problem. There are similar works that impose variational constraints on random point samples for images \cite{variational}, as well as multi-resolution techniques for volumetric data \cite{onvari} and an extension to box spline spaces \cite{xu2012rec}. In the same vein, we seek a function $g \in \aspace{\lattice{L}_h}{\varphi}$ that attempts to interpolate the input data, and minimize a functional $E$. Specifically, we seek a $g$ that minimizes {\small 
\begin{equation} \label{eq:minexp}
 	\sump_{i=1}^{M} (g(\vx_i) -  f_i)^2 + E(g).
\end{equation}}
It is important to introduce a bit of notation before proceeding, $\dv{\mathbf{v}} f := \mathbf{v} \cdot \grad f$ defines the directional derivative of $f$ in the direction $\mathbf{v}$. The Beppo-Levi inner product of order $n$ is defined as $\blip{n}{f}{g}:= \sum_{|\mathbf{p}|=n}({n!}/{\mathbf{p}!}) \innerp{\mathbf{D^p}f}{\mathbf{D^p}g}$, where $\mathbf{p}$ is an integer vector such that $\left| \mathbf{p}\right| := p_1 + p_2 + p_3 $. We also use the short hand $\mathbf{D^p}f := \partial^{\left| \mathbf{p}\right|}f/({\partial^{p_1}x_1}{\partial^{p_2}x_2}{\partial^{p_3}x_3}).$ The second order Beppo-Levi norm of $f$ is defined as $\bln{2}{f}^2:=\blip{2}{f}{f}$.

@@

Finding $g$ that minimizes equation \EQ{eq:minexp} involves rewriting the minimization problem in terms of $\mathbf{c}$ \cite{xu2012rec}. Following some straightforward algebraic manipulations (see \cite{xu2012rec} for details), the coefficient vector can be obtained by taking $\mathbf{c} = ( \mathbf{\Phii}^T\mathbf{\Phii} + \lambda_1\mathbf{G} + \lambda_2\mathbf{S})^{-1}\mathbf{\Phii}^T\mathbf{f}$, where $G_{i,j}:=\innerp{\varphi_i}{\varphi_j}$, $S_{i,j}:=\blip{2}{\varphi_i}{\varphi_j}$, and $P_{i,j} := \varphi_j(\vx_i)$ as before. Since $\varphi$ is known in closed form in our test cases, the entries of both $\mathbf{G}$ and $\mathbf{S}$ can be evaluated analytically. The above vector equation can be solved using the Conjugate Gradient method.

### 2015-04-08 results.tex
Bringing everything up to date

\section{Experiments}

@@

\subsubsection{Basis function}
Choosing $\mathbf{\Xii = \Xii_{C2} := \left[ I_3  \ I_3 \right]}$ (the $3 \times 6$ matrix created by the union of two identity matrices) generates a triliniear piecewise polynomial. It is easy to see, by definition, when $\mathbf{\Xii_{C2}}$ is used in conjunction with \EQp{eq:box_def}, the resulting function $M_\mathbf{\Xii_{C2}}(\vx)$ corresponds to the non-centered tensor product extension of the linear B-spline in three dimensions. Likewise, the non-centered cubic tensor product box spline is given by $M_\mathbf{\Xii_{C4}}(\vx)$, where $\mathbf{\Xii_{C4} := \left[ I_3  \ I_3 \ I_3  \ I_3 \right]}$. A box spline is easily centered by taking $M_\mathbf{\Xii}(\vx + \mathbf{c_\Xii})$, with $\mathbf{c_\Xii}:=\sum_{\xi \in \mathbf{\Xii}} \frac{1}{2}\xi$. We will assume, unless otherwise stated, that $M_\mathbf{\Xii}$ defines the centered box spline generated by $\mathbf{\Xii}$.

@@

generates a piecewise linear polynomial function \cite{practicalbox}. The support of this box spline is a rhombic dodecahedron, which incidentally contains the set of first nearest neighbors to the Voronoi region of the lattice at the origin, and is a natural fit for the lattice. The quintic box spline is given by the matrix $\mathbf{\Xii_{B2}} := \left[\mathbf{\Xii_{B}} \ \mathbf{\Xii_{B}}\right]$, or may also be seen as the convolution of the linear box spline with itself.  However, there exist both an efficient piecewise polynomial representation \cite{practicalbox}, and an efficient implementation \cite{fastbox} of these piecewise quintic polynomial functions.

>>>>

\section{Experiments}
To test our methodology, we choose comparable basis functions on both the Cartesian and BCC lattices. These functions are comparable in that they both provide the same approximation order and smoothness.

@@

\subsubsection{Basis function}
Choosing $\mathbf{\Xii = \Xii_{C2} := \left[ I_3  \ I_3 \right]}$ (the $3 \times 6$ matrix created by the union of two identity matrices) generates a triliniear piecewise polynomial. It is easy to see, by definition, when $\mathbf{\Xii_{C2}}$ is used in conjunction with \EQp{eq:box_def}, the resulting function $M_\mathbf{\Xii_{C2}}(\vx)$ corresponds to the non-centered tensor product extension of the linear B-spline in three dimensions. Likewise, the non-centered cubic tensor product box spline is given by $M_\mathbf{\Xii_{C4}}(\vx)$, where $\mathbf{\Xii_{C4} := \left[ \Xii_{C2} \ \Xii_{C2} \right]}$. A box spline is easily centered by taking $M_\mathbf{\Xii}(\vx + \mathbf{c_\Xii})$, with $\mathbf{c_\Xii}:=\sum_{\xi \in \mathbf{\Xii}} \frac{1}{2}\xi$. We will assume, unless otherwise stated, that $M_\mathbf{\Xii}$ defines the centered box spline generated by $\mathbf{\Xii}$. Due to the smoothness requirements of the optimization constraint, we only considered the spline $M_{\Xii_{C4}}$.

\subsubsection{Filter weights}
This is where we derive filter weights for the experiments.

@@

generates a piecewise linear polynomial function \cite{practicalbox}. The support of this box spline is a rhombic dodecahedron, which incidentally contains the set of first nearest neighbors to the Voronoi region of the lattice at the origin, and is a natural fit for the lattice. The quintic box spline is given by the matrix $\mathbf{\Xii_{B2}} := \left[\mathbf{\Xii_{B}} \ \mathbf{\Xii_{B}}\right]$, or may also be seen as the convolution of the linear box spline with itself.  However, there exist both an efficient piecewise polynomial representation \cite{practicalbox}, and an efficient implementation \cite{fastbox} of these piecewise quintic polynomial functions. Again, due to the smoothness requirements of the optimization constraint, we only considered the spline $M_{\Xii_{B2}}$.
